BRiM: A Modular Bicycle-Rider Modeling Framework

### 2024-01-16 paper.tex

The development of computationally efficient and validated single-track vehicle-rider models has traditionally required handcrafted one-off models. Here we introduce \brim{}, a software package that facilitates building these models in a modular fashion while retaining access to the mathematical elements for handcrafted modeling when desired. We demonstrate the flexibility of the software by constructing the Carvallo-Whipple bicycle model with different numerical parameters representing different bicycles, modifying it with a front fork suspension travel model, and extending it with moving rider arms driven by joint torques at the elbows. Using these models we solve a lane-change optimal control problem for six different model variations which solve in mere seconds on a modern laptop. Our tool enables flexible and rapid modeling of single-track vehicle-rider models that give precise results at high computational efficiency.

@@

A \term{connection} is a utility of a parent to define relationships between two or more child submodels. Connections provide three key features.
Firstly, connections enable parent models to define the interactions between their child submodels in a modular fashion.
Secondly, connections enable the reuse of these possibly complex interactions.
An example of both is the interaction between a wheel and the ground, where multiple complex descriptions of tire models are possible.
Using connections these tire models can be defined once and then be reused in all bicycle models in \brim{}.
Lastly, connections allow new properties to be incorporated into a model.
For example, when modeling a bicycle with a leaning rider, the axis should be defined about which the rider leans.
However, if a bicycle is being modeled without a rider then this axis is not needed and so not considered.

@@

While it is possible to implement any bicycle model using \brim{}, \brim{} distributes a complete implementation of the Carvallo-Whipple bicycle model~\citep{carvallo_1901_theorie,whipple_1899_stability} following the parameterization convention of \citet{moore_2012_human}.
The Carvallo-Whipple bicycle is widely recognized as the lowest-order bicycle model to have been repeatedly validated by multiple authors~\citep{kooijman_2008_experimental,schwab_2013_review,sharp_2008_stability}.
The parametrization convention of \citeauthor{moore_2012_human}, as shown in Figure~\ref{fig:moore-whipple-bicycle}, has been chosen due to its configuration independence, which is preferred when forming nonlinear EoMs of a bicycle,  \citep{peterson_2013_bicycle}.
By default, the Carvallo-Whipple bicycle consists of a ground and four bodies: a rear wheel, a rear frame, a front frame, and a front wheel.
Each body is assumed to be rigid and all bodies are attached using pin joints.
The rear wheel's contact point is defined within the ground plane and the rear frame is oriented in yaw-roll-pitch rotation relative to the ground.
The front wheel's contact point is constrained to the ground using a holonomic constraint, and nonholonomic constraints are applied to both wheels to enforce no-slip conditions.
Further details of the Carvallo-Whipple bicycle model's definition are found in \citet{moore_2012_human}.

@@

The next three are dynamic differential equations, each one corresponding to the time derivative of the three independent generalized speeds
\begin{equation}
    \mathbf{M} \dot{\mathbf{u}}_{ind} - \mathbf{k} = \mathbf{0} \, ,
\end{equation}
where $\mathbf{M}$ is the $3 \times 3$ mass matrix of the bicycle model and $\mathbf{k}$ is its length-3 forcing column vector.

@@

In the trajectory tracking optimization problems, the task is to follow a target path representing a lane change.
The path is defined by the equation
\begin{equation}
    \bar{q}_2\left(q_1\right) = \begin{cases}
        0 \, , & \hphantom{X_1 - X_S \leq } q_1 < X_s \\ 
        \frac{1}{2} X_2\left(1 - \cos{\left(\frac{q_1 - X_s}{X_1 - 2X_s}\pi\right)}\right) \, , & \hphantom{X_1 - {}} X_s \leq q_1 \leq X_1 - X_s \\
        X_2 \, , & X_1 - X_s < q_1 \\
    \end{cases} \, ,
\end{equation}
where $q_1$ and $q_2$ are the states describing the position of the bicycle in the $x$ (longitudinal) and $y$ (lateral) directions of the Newtonian frame respectively, $\bar{q}_2$ is the target lateral position as a function of $q_1$, $X_1$ and $X_2$ are the longitudinal length and lateral displacement of the target path respectively, and $X_s$ is the length of straight at the beginning and end of the path.

The objective function $J$ simultaneously minimizes both the tracking error and the inputs.
It is defined as
\begin{equation}
    J = \int_{t_0}^{t_F} (1-w) \left(q_2 - \bar{q}_2\left(q_1\right) \right)^2 dt  + w \sum_{i=1}^{R} \int_{t_0}^{t_F} r_i^2 dt \, ,
\end{equation}
where $t_0$ and $t_F$ are the initial and final times, $r_i$ is the $i$th input variable, and $w$ is a weighting parameter between the squared tracking error and the sum of squared inputs.
We selected $w=10^{-4}$, which aims to target a mean tracking error of \SI{0.025}{\meter} for an approximate torque of \SI{2.5}{\newton\meter}.

We add constraints defining the initial and final state of the system's motion to be in the nominal configuration.
In optimizations \#\numrange[range-phrase=--]{1}{6} we enforce
\begin{equation*}
    \begin{aligned}
        q_1\left(t_0\right) &= 0.0 \\
        q_2\left(t_0\right) &= 0.0 \\
        q_3\left(t_0\right) &= 0.0 \\
        q_4\left(t_0\right) &= 0.0
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        q_1\left(t_F\right) - X_1 &= 0.0 \\
        q_2\left(t_F\right) - X_2 &= 0.0 \\
        q_3\left(t_F\right) &= 0.0 \\
        q_4\left(t_F\right) &= 0.0
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        q_6\left(t_0\right) &= 0.0 \\
        q_7\left(t_0\right) &= 0.0 \\
        q_8\left(t_0\right) &= 0.0 \\
        \\
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        \\
        q_7\left(t_F\right) &= 0.0 \\
        \\
        \\
    \end{aligned}
\end{equation*}

The velocity at the end of the path is enforced to be the same as at the start using periodic constraints on the angular velocity of both wheels, $u_6$ and $u_8$.
Additionally, in optimization \#4 the generalized coordinate and generalized speed for the suspension are constrained at $t_0$ such that the suspension compression and weight through the handlebars are in equilibrium.

@@

The initial guess in each optimization was created using a forward simulation that aimed to simultaneously satisfy $\tilde{q}_1\left(t_0\right)=0$, $\tilde{q}_2\left(t_0\right)=0$, $\tilde{q}_1\left(t_F\right)=X_1$, and $\tilde{q}_2\left(t_F\right)=X_2$.
The tilde above a variable denotes that this is the initial guess corresponding to that variable.
The forward simulation from $t_0$ until $t_F$ was conducted under null inputs ($\tilde{r}_i = 0, i \in \mathbb{N} \cap \left[1,R\right]$).
Due to the null inputs, the initial state required modification such that the bicycle rides diagonally between the endpoints:
\begin{equation*}
    \tilde{q}_3\left(t_0\right)=\atantwo\left(\frac{X_2}{X_1}\right) \qquad \qquad \tilde{u}_4\left(t_0\right)=0 \qquad \qquad \tilde{u}_6\left(t_0\right)=-\frac{\sqrt{X_1^2 + X_2^2}}{r_R\left(t_F - t_0\right)} \qquad \qquad \tilde{u}_7\left(t_0\right)=0 \, .
\end{equation*}

As the system dynamics are defined as differential algebraic equations, the \code{IDA} differential algebraic equation solver~\citep{gardner_2022_sundials,hindmarsh_2005_sundials} from the \softwarepackage{ODES} Scikit~\citep{malengier_2018_odes} was used to conduct the forward simulation.
The resulting initial guesses are dynamically feasible, with the only active instance constraint violations being
\begin{equation*}
    q_3\left(t_0\right) = \atantwo\left(\frac{X_2}{X_1}\right) \neq 0 \qquad \qquad
    q_3\left(t_F\right) = \atantwo\left(\frac{X_2}{X_1}\right) \neq 0 \, .
\end{equation*}

@@

The optimal controls are shown in Figure~\ref{fig:torques_all}.
Comparisons are shown for the optimal propulsion/braking torque for all optimizations and optimal steering torque for optimizations \#\numrange[range-phrase=--]{1}{5}.
For optimization \#6, where there was no steering torque, the optimal left and right elbow torques are included.
A final subplots compares the tracking error of the optimal trajectory in each problem.

@@

\begin{table}
\centering
\begin{tabular}{cSSSSSS}
\toprule
Optimization & {Optimal cost}             & {Tracking cost}         & {Input cost} & {Mean tracking error (\si{\meter})} & {\# NLP iterations} & {Time in \ipopt{} (\si{\second})} \\
\midrule
\#1          & \num{3.723e-3} & \num{2.832e-3} & 8.909    & 0.038             & 45             & 3.1                               \\
\#2          & \num{1.959e-3} & \num{1.388e-3} & 5.718    & 0.026             & 39             & 2.5                               \\
\#3          & \num{1.925e-3} & \num{1.324e-3} & 6.114    & 0.026             & 56             & 3.4                              \\
\#4          & \num{2.207e-3} & \num{1.661e-3} & 5.463    & 0.029             & 99             & 9.7                              \\
\#5          & \num{7.726e-3} & \num{5.723e-3} & 20.040   & 0.053             & 136            & 30.0                              \\
\#6          & \num{8.630e-3} & \num{5.978e-3} & 26.530   & 0.055             & 160            & 29.0    \\   \bottomrule
\end{tabular}
\caption{Solution and convergence metrics for optimization \#\numrange[range-phrase=--]{1}{6}. Computational metrics were produced using an \textit{Intel(R) Core(TM) i7-8750H CPU @ 2.20GHz}.}
\label{tab:results}
\end{table}

@@

In the optimal controls for all six problems, the steering torque near $t_0$ acts to roll the bicycle toward the centerline of the target path while steering the bicycle away from it.
This is to be expected as the mass center must be inside the wheels during a turn for the centripetal forces to balance the bicycle.
The optimal controls are much smaller in magnitude after this initial extreme torque.

All optimal trajectories generally cross the target path in the middle.
This can be explained by the target path's rotational symmetry and the periodicity constraints imposed at $t_0$ and $t_F$ in all of the optimization problems.
The increased tracking error and input cost of optimization \#1 compared to the other riderless optimization \#\numrange[range-phrase=--]{2}{4} can be explained by the lower agility of the Batavus Browser bicycle.
It is designed to be a stable city bicycle and is almost twice as heavy as the Gary Fisher and Bianchi Pista bicycles.

The solutions to optimizations \#3 and \#4 are similar by multiple measures. Qualitatively, the optimal states (Figure~\ref{fig:states_all}) and inputs (Figure~\ref{fig:torques_all}) appear near identical. Quantitatively, the optimal cost is 14.6\% larger for optimization \#4 due to a larger tracking error, despite a lower input cost. As the target trajectory follows a path on flat ground, it should be expected that there will be very little to no travel in the suspension during the maneuver, and therefore the presence of suspension should have little effect on the results.

Comparing the solutions of optimizations \#1 and \#5, the shape of the optimal trajectories differs with much larger deviations from the target path at the transitions between the straights and the chicane being exhibited in optimization \#5. This is to be expected as the addition of the rider results in a large increase in the mass and inertia of the system, and an increase in the height of the mass center. The optimal cost is more than twice as large in optimization \#5, which can be attributed in equal parts to both more-than-doubling of the tracking and input costs. Despite the initial input from the steering torque being smaller, the steering torque during optimization \#5 remains at a much higher level during the maneuver, which is required to control the additional mass and inertia of the rider.

@@

All optimization problems were solved using \opty{}'s backwards Euler collocation.
With this method, the state within a given mesh interval is only affected by the inputs at the interval final time.
In other words, the inputs only influence the states in the preceding mesh section.
Therefore, the inputs at $t_0$ do not influence the states at all.
However, the inputs at $t_0$ do contribute to the objective function.
In all optimizations, the inputs at $t_0$ are always optimized to 0.0 because this is the value that minimizes the objective function and doing so does not affect the dynamics in any way.
This is a limitation of the backwards Euler method and the results could be improved if a discretization scheme that collocates the initial time was used.

@@

A limitation of \brim{} is that it forces model developers to use an existing interface for augmented model components. These interfaces can limit flexibility, especially if a component requires additional properties to be interacted with.
While connections largely solve this problem, an over-usage of them can result in a cluttering of the codebase.

>>>>

The development of computationally efficient and validated single-track vehicle-rider models has traditionally required handcrafted one-off models. Here we introduce \brim{}, a software package that facilitates building these models in a modular fashion while retaining access to the mathematical elements for handcrafted modeling when desired. We demonstrate the flexibility of the software by constructing the Carvallo-Whipple bicycle model with different numerical parameters representing different bicycles, modifying it with a front fork suspension travel model, and extending it with moving rider arms driven by joint torques at the elbows. Using these models we solve a lane-change optimal control problem for six different model variations which solve in mere seconds on a modern computer. Our tool enables flexible and rapid modeling of single-track vehicle-rider models that give precise results at high computational efficiency.

@@

A \term{connection} is a utility of a parent model to define relationships between two or more child submodels. Connections provide three key features.
Firstly, connections enable parent models to define the interactions between their child submodels in a modular fashion.
Secondly, connections enable the reuse of these possibly complex interactions.
An example of both is the interaction between a wheel and the ground, where multiple complex descriptions of tire models are possible.
Using connections these tire models can be defined once and then be reused in all bicycle models in \brim{}.
Lastly, connections allow new properties to be incorporated into a model.
For example, when modeling a bicycle with a leaning rider, the axis should be defined about which the rider leans.
However, if a bicycle is being modeled without a rider then this axis is not needed and so not considered.

@@

While it is possible to implement any bicycle model using \brim{}, \brim{} distributes a complete implementation of the Carvallo-Whipple bicycle model~\citep{carvallo_1899_theorie,whipple_1899_stability} following the parameterization convention of \citet{moore_2012_human}.
The Carvallo-Whipple bicycle is widely recognized as the lowest-order bicycle model to have been repeatedly validated by multiple authors~\citep{kooijman_2008_experimental,schwab_2013_review,sharp_2008_stability}.
The parametrization convention of \citeauthor{moore_2012_human}, as shown in Figure~\ref{fig:moore-whipple-bicycle}, has been chosen due to its configuration independence, which is preferred when forming nonlinear EoMs of a bicycle,  \citep{peterson_2013_bicycle}.
By default, the Carvallo-Whipple bicycle consists of a ground and four bodies: a rear wheel, a rear frame, a front frame, and a front wheel.
Each body is assumed to be rigid and all bodies are attached using pin joints.
The rear wheel's contact point is defined within the ground plane and the rear frame is oriented in yaw-roll-pitch rotation relative to the ground.
The front wheel's contact point is constrained to the ground using a holonomic constraint, and nonholonomic constraints are applied to both wheels to enforce no-slip conditions.
Further details of the Carvallo-Whipple bicycle model's definition are found in \citet{moore_2012_human}.

@@

The next three are dynamic differential equations, each one corresponding to the time derivative of the three independent generalized speeds
\begin{equation}
    \mathbf{M} \dot{\mathbf{u}}_{ind} - \mathbf{k} = \mathbf{0} \, ,
\end{equation}
where $\mathbf{M}$ is the $U \times U$ mass matrix of the bicycle model and $\mathbf{k}$ is its length-U forcing column vector encompassing both externally applied forces and forces due to various velocity effects.

@@

In the trajectory tracking optimization problems, the task is to follow a target ground path representing a lane change.
The path is defined by the equation
\begin{equation}
    \bar{q}_2\left(q_1\right) = \begin{cases}
        0 \, , & \hphantom{X_1 - X_S \leq } q_1 < X_s \\ 
        \frac{1}{2} X_2\left(1 - \cos{\left(\frac{q_1 - X_s}{X_1 - 2X_s}\pi\right)}\right) \, , & \hphantom{X_1 - {}} X_s \leq q_1 \leq X_1 - X_s \\
        X_2 \, , & X_1 - X_s < q_1 \\
    \end{cases} \, ,
\end{equation}
where $q_1$ and $q_2$ are the states describing the $x$ (longitudinal) and $y$ (lateral) position of the rear wheel contact point in the Newtonian coordinate system respectively, $\bar{q}_2$ is the target lateral position as a function of $q_1$, $X_1$ and $X_2$ are the longitudinal length and lateral displacement of the target path respectively, and $X_s$ is the length of straight at the beginning and end of the path.

The objective function $J$ simultaneously minimizes both the tracking error and the input magnitudes, to find controls that minimize the approximated energy used in the maneuver.
It is defined as
\begin{equation}
    J = \int_{t_0}^{t_F} (1-w) \left(q_2 - \bar{q}_2\left(q_1\right) \right)^2 dt  + w \sum_{i=1}^{R} \int_{t_0}^{t_F} r_i^2 dt \, ,
\end{equation}
where $t_0$ and $t_F$ are the initial and final times, $r_i$ is the $i$th input variable, and $w$ is a weighting parameter between the squared tracking error and the sum of squared inputs.
We selected $w=2.5 \cdot 10^{-3}$, which aims to target a mean tracking error of \SI{0.025}{\meter} for an approximate torque of \SI{0.5}{\newton\meter}.

We add constraints defining the initial and final state of the system's motion to be in the nominal configuration and at a nominal travel speed.
In optimizations \#\numrange[range-phrase=--]{1}{6} we enforce
\begin{equation*}
    \begin{aligned}
        q_1\left(t_1\right) &= 0.0 \\
        q_2\left(t_1\right) &= 0.0 \\
        q_3\left(t_1\right) &= 0.0 \\
        q_4\left(t_1\right) &= 0.0
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        q_1\left(t_F\right) - X_1 &= 0.0 \\
        q_2\left(t_F\right) - X_2 &= 0.0 \\
        q_3\left(t_F\right) &= 0.0 \\
        q_4\left(t_F\right) &= 0.0
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        q_6\left(t_1\right) &= 0.0 \\
        q_7\left(t_1\right) &= 0.0 \\
        q_8\left(t_1\right) &= 0.0 \\
        \\
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        \\
        q_7\left(t_F\right) &= 0.0 \\
        \\
        \\
    \end{aligned}
\end{equation*}

where the second node at $t_1$ is used for the initial state because the inputs do not affect the first node when using backward Euler discretization.
% The velocity at the end of the path is enforced to be the same as at the start using periodic constraints on the angular velocity of the rear wheel, $u_6$.
Additionally, in optimization \#4 the generalized coordinate and generalized speed for the suspension are constrained at $t_0$ such that the suspension compression and weight through the handlebars are in equilibrium.

@@

The initial guess in each optimization was created using a forward simulation that aimed to simultaneously satisfy $\tilde{q}_1\left(t_0\right)=0$, $\tilde{q}_2\left(t_0\right)=0$, $\tilde{q}_1\left(t_F\right)=X_1$, and $\tilde{q}_2\left(t_F\right)=X_2$.
The tilde above a variable denotes that this is the initial guess corresponding to that variable.
The forward simulation from $t_0$ until $t_F$ was conducted under null inputs ($\tilde{r}_i = 0, i \in \mathbb{N} \cap \left[1,R\right]$).
Due to the null inputs, the initial state required modification such that the bicycle rides diagonally between the endpoints:
\begin{equation*}
    \tilde{q}_3\left(t_0\right)=\arctan\left(\frac{X_2}{X_1}\right) \qquad \qquad \tilde{u}_4\left(t_0\right)=0 \qquad \qquad \tilde{u}_6\left(t_0\right)=-\frac{\sqrt{X_1^2 + X_2^2}}{r_R\left(t_F - t_0\right)} \qquad \qquad \tilde{u}_7\left(t_0\right)=0 \, .
\end{equation*}

As the system dynamics are defined as differential algebraic equations, the \code{IDA} differential algebraic equation solver~\citep{gardner_2022_sundials,hindmarsh_2005_sundials} from the \softwarepackage{ODES} Scikit~\citep{malengier_2018_odes} was used to conduct the forward simulation.
The resulting initial guesses are dynamically feasible, with the only active instance constraint violations being
\begin{equation*}
    q_3\left(t_0\right) = \arctan\left(\frac{X_2}{X_1}\right) \neq 0 \qquad \qquad
    q_3\left(t_F\right) = \arctan\left(\frac{X_2}{X_1}\right) \neq 0 \, .
\end{equation*}

@@

The optimal controls are shown in Figure~\ref{fig:torques_all}.
Comparisons are shown for the optimal propulsion/braking torque for all optimizations and optimal steering torque for optimizations \#\numrange[range-phrase=--]{1}{5}.
For optimization \#6, where there was no steering torque, the optimal left and right elbow torques are included.
A final subplot compares the tracking error of the optimal trajectory in each problem.

@@

\begin{table}
\centering
\begin{tabular}{cSSSSSS}
\toprule
Optimization & {Optimal cost}             & {Tracking cost}         & {Input cost} & {Mean tracking error (\si{\meter})} & {\# NLP iterations} & {Time in \ipopt{} (\si{\second})} \\
\midrule
\#1          & \num{1.798e-3} & \num{4.971e-4} & \num{5.229e-1}    & \num{1.568e-2}             & 17             & 0.3                               \\
\#2          & \num{1.622e-3} & \num{1.085e-4} & \num{6.056e-1}    & \num{7.366e-3}             & 26             & 0.4                               \\
\#3          & \num{1.461e-3} & \num{1.312e-4} & \num{5.320e-1}    & \num{8.101e-3}             & 22             & 0.4                              \\
\#4          & \num{1.472e-3} & \num{1.316e-4} & \num{5.363e-1}    & \num{8.112e-3}             & 51             & 1.3                              \\
\#5          & \num{1.429e-2} & \num{6.544E-3} & \num{3.106e0}   & \num{5.720e-2}             & 13            & 0.7                              \\
\#6          & \num{1.848e-2} & \num{9.431E-3} & \num{3.630e0}   & \num{6.867e-2}             & 13            & 0.6    \\   \bottomrule
\end{tabular}
\caption{Solution and convergence metrics for optimization \#\numrange[range-phrase=--]{1}{6}. Computational metrics were produced using an \textit{Intel(R) Core(TM) i9-13900K CPU @ 3.00GHz}.}
\label{tab:results}
\end{table}

@@

In the optimal controls for all six problems, countersteering near $t_0$ acts to roll the bicycle toward the centerline of the target path while steering the bicycle away from it.
This is to be expected as the mass center must be inside the wheels during a turn for the centripetal forces to balance the bicycle.

All optimal trajectories generally cross the target path in the middle.
This can be explained by the target path's rotational symmetry and the periodicity constraints imposed at $t_0$ and $t_F$ in all of the optimization problems.
The propulsion/braking torque remains low in each optimization, as the path can mostly be followed without braking.
The increased tracking error of optimization \#1 compared to the other riderless optimization \#\numrange[range-phrase=--]{2}{4} can be explained by the lower agility of the Batavus Browser bicycle.
It is designed to be a stable city bicycle and is almost twice as heavy as the Gary Fisher and Bianchi Pista bicycles.

The solutions to optimizations \#3 and \#4 are similar by multiple measures. Qualitatively, the optimal states (Figure~\ref{fig:states_all}) and inputs (Figure~\ref{fig:torques_all}) appear nearly identical. As the target trajectory follows a path on flat ground, it should be expected that there will be very little to no travel in the suspension during the maneuver, and therefore the presence of suspension should have little effect on the results.

Comparing the solutions of optimizations \#1 and \#5, the shape of the optimal trajectories differs with much larger deviations from the target path at the transitions between the straights and the chicane being exhibited in optimization \#5. This is to be expected as the addition of the rider results in a large increase in the mass and inertia of the system, and an increase in the height of the mass center. The optimal cost is approximately eight times as large in optimization \#5, which can be attributed in equal parts to both the tracking and input costs.

@@

All optimization problems were solved using \opty{}'s backward Euler discretization.
With this method, the state within a given mesh interval is only affected by the inputs at the interval final time.
In other words, the inputs only influence the states in the preceding mesh section.
Therefore, the inputs at $t_0$ do not influence the states at all.
However, the inputs at $t_0$ do contribute to the objective function.
In all optimizations, the inputs at $t_0$ are always optimized to 0.0 because this is the value that minimizes the objective function and doing so does not affect the dynamics in any way.
This is a limitation of the backward Euler method and the results could be improved if a discretization scheme that collocates the initial time was used.

@@

A limitation of \brim{} is that it forces model developers to use an existing interface for augmented model components. These interfaces can limit flexibility, especially if a component requires additional properties to be interacted with.
While connections largely solve this problem, an over-usage of them can result in a cluttering of the implementation code.
